---
title: "within subjects speaker reliability manipulation"
author: "jdegen"
date: "February 24, 2015"
output: html_document
---

```{r}
load("results/data/r.RData")
```

Normalized mean responses by slider (column), quantifier (row), and trial type (color).

```{r, echo=FALSE, warning=F, message=FALSE}
agrr = aggregate(normresponse ~ PriorExpectation + Proportion + quantifier + Item + trial_type,data=r,FUN=mean)
#agrr = droplevels(subset(agrr, Proportion == "100"))
ggplot(agrr, aes(x=PriorExpectation, y=normresponse, color=trial_type)) +
  geom_point() +
  #geom_errorbar(aes(ymin=YMin,ymax=YMax)) +
  geom_smooth() +
  facet_grid(quantifier~Proportion,scales="free_y") +
  ylab("Mean normalized response") 

```

Normalized mean responses on 100%/all slider by block (column), quantifier (row), and trial type (color). Drunk/court are unreliable/uncooperative, sober is reliable. Clear effect of speaker reliability on "some", but values still not super high (normalization issue).

```{r, echo=FALSE, warning=F, message=FALSE}

agrr = aggregate(normresponse ~ PriorExpectation + Proportion + quantifier + Item + trial_type + block,data=r,FUN=mean)
agrr = droplevels(subset(agrr, Proportion == "100"))
ggplot(agrr, aes(x=PriorExpectation, y=normresponse, color=trial_type)) +
  geom_point() +
  geom_smooth() +
  facet_grid(quantifier~block,scales="free_y") +
  ylab("Mean normalized response")

```

Raw mean responses on 100%/all slider by block (column), quantifier (row), and trial type (color). Drunk/court are unreliable/uncooperative, sober is reliable. Very clear effect of speaker reliability on "some", and for the first time, values approaching 1 in the unreliable case. This has never happened before!


```{r, echo=FALSE, warning=F, message=FALSE}

agrr = aggregate(response ~ PriorExpectation + Proportion + quantifier + Item + trial_type + block,data=r,FUN=mean)
agrr = droplevels(subset(agrr, Proportion == "100"))
ggplot(agrr, aes(x=PriorExpectation, y=response, color=trial_type)) +
  geom_point() +
  geom_smooth() +
  facet_grid(quantifier~block,scales="free_y") +
  ylab("Mean raw response")

```

