---
title: "wonkyworlds"
author: "mht"
date: "October 15, 2015"
output: html_document
---


```{r helpers, echo=FALSE}
library(coda)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10)   


reconstructHist <- function(df, samples){
  Map(function(x){
    
  },names(df)
}

```

# Prior data

First, load the prior data, primarily to create the x-axis for our plots.


## Load human subjects data, already histogrammed and smoothed


```{r}
priorprobs = read.table(file="~/Documents/research/sinking-marbles/experiments/24_sinking-marbles-prior-fourstep/results/data/smoothed_15marbles_priors_withnames.txt",sep="\t", header=T, quote="")

# prior expectation for each item
prior_exps <- priorprobs %>%
  gather(State,Probability,X0:X15) %>%
  mutate(State = as.numeric(as.character(gsub("X","",State)))) %>%
  group_by(Item) %>%
  summarise(exp.val=sum(State*Probability))
  

# prior all-state probability for each item
prior_exps = left_join(prior_exps,
                           priorprobs[,c("Item","X15")])

qplot(data=prior_exps, x = exp.val, geom='histogram')+
  ggtitle("Distribution of items' prior expected values")+
  xlab("Expected Value of P(x), smoothed empirical")
```

## Load results of prior (BDA) model of prior elicitation data

```{r}

path.prior <- "~/Documents/research/sinking-marbles/bayesian_model_comparison/priors/results/"

mhiter <- 75000
burn <- mhiter/2
samples <- mhiter - burn

#prefix <- "prior4stepBDA-binomials-fullPosterior_incrMH"
#prefix <- "priorBDA-binomials_incrMH"
#prefix <- "priorBDA-mix3binomials-fullPosterior_incrMH"
prefix <- "priorBDA-mix2binomials-fullPosterior_incrMH"
#prefix <- "priorBDA-mix2binomials-phi-removePhi-fullPosterior_incrMH"

d0<-read.csv(paste(path.prior, prefix, mhiter, "burn",  burn, ".csv", sep=""))


# backout samples from webppl's "histogram" structure
d.thetas <- data.frame(Parameter = rep(d0$Parameter, 1+samples*d0$Probability),
                       Item = rep(d0$Item, 1+samples*d0$Probability),
                       Response = rep(to.n(d0$Value), 1+samples*d0$Probability))


## when you have data by bin
## compute posterior predictive of prior data
# grab parameters of interest (Items, Bins), compute MAP for each Item x Bin
inferred.priors <- d.thetas %>% 
  filter(!(Parameter%in% 
             c('theta',"theta1","theta2","theta3",
               "w1","w2","w3",
               'mix','phi'))) %>% 
  group_by(Item, Parameter) %>%
  summarise(MAP = estimate_mode(Response)) %>%
  ungroup() %>%
  rename(bins=Parameter) %>%
  mutate(bins = to.n(bins)) %>%
  group_by(Item) %>%
  summarise(exp.val = sum(bins*MAP),
            allprob =sum ((bins==15)*MAP))


qplot(data=inferred.priors, x= exp.val, geom = 'histogram')+
  ggtitle("Distribution of items' prior expected values")+
  xlab("Expected Value of P(x), posterior predictive of prior model")

range(inferred.priors$exp.val)
```

## Compare data and model

Compare prior elicitation data (smoothed, binned) to posterior predictive of the BDA model.

Right now, only compare expected values

```{r}
priors.md<- left_join(inferred.priors, prior_exps %>% rename(emp.exp.val = exp.val))

with(priors.md, cor(exp.val, emp.exp.val))^2
with(priors.md, cor(allprob, X15))^2
```


# Posterior (Interpretation) data

## Load human subjects data
