---
title: "wonky-priors-bda"
author: "mht"
date: "September 21, 2015"
output: html_document
---

```{r helpers}

library(coda)
library(hydroGOF)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10)   
```

Load prior data, convert histogram from wppl into samples, calculate credible intervals, plot


```{r}

mhiter <- 75000
burn <- mhiter/2
samples <- mhiter - burn
fpath<- "~/Documents/research/sinking-marbles/bayesian_model_comparison/priors/results/"
#prefix <- "priorBDA-binomials_incrMH"
#prefix <- "prior4stepBDA-binomials-fullPosterior_incrMH"
#prefix <- "priorBDA-binomials-fullPosterior-phi_incrMH"
#prefix <- "priorBDA-mix2binomials-phi-removePhi-fullPosterior_incrMH"
prefix <- "priorBDA-mix2binomials-fullPosterior_incrMH"

#fpath<- "~/Documents/research/sinking-marbles/bayesian_model_comparison/results/"
#prefix<-"regularRSA_fullPosterior_unifBackoff_justCompState_wRepData_bdaPriors2thetas50k_so_wp_phi_incrMH"
#prefix <- "priorBDA-binomials-byItemPhi_incrMH"

d0<-read.csv(paste(fpath, prefix, mhiter, "burn",  burn, ".csv", sep=""))

#d.prior<-d0 %>% filter(Parameter=='prior')


d.params <- data.frame(Parameter = rep(d0$Parameter, 
                                         1+samples*d0$Probability),
                         Item = rep(d0$Item, 
                                    1+samples*d0$Probability),
                         Response = rep(d0$Value, 
                                        1+samples*d0$Probability))



d.tidy <- d.params %>% 
  filter(Parameter%in%c('theta', "theta1", "theta2", "theta3",
                                             "mix", "w1", "w2", "w3")) %>% 
  group_by(Parameter,Item) %>%
  summarise(MAP = estimate_mode(Response),
            credLow = HPDlo(Response),
            credHigh = HPDhi(Response))

ggplot(data=d.tidy %>% filter(Parameter=='mix'), 
      aes(x = MAP) 
      )+
  geom_histogram(binwidth=0.02)+
  ggtitle('mixture component')

plotname <- paste("mixtureHistogram-",prefix, mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath,"figures/", plotname, sep=''), 
       width = 8, height = 5)


```


"Bi-modality score" = || theta1 - theta2 || * 2* (0.5-|| mix-0.5 ||)

```{r}
d.biModal<-d.tidy %>% 
  select(Parameter, Item, MAP) %>%
  spread(Parameter, MAP) %>%
  mutate(absMix = 2 * (0.5 - abs(mix-0.5)),
         diffTheta = abs(theta1-theta2),
         bimodality = abs(theta1-theta2) * 2 * (0.5 - abs(mix-0.5)),
         Item = reorder(Item, bimodality, function(x) x)) %>%
    select(-theta1, -theta2, -mix) %>%
    gather(param, val, -Item)


ggplot(d.biModal, aes(x=Item, y = val))+
  geom_bar(stat='identity', position=position_dodge())+
  coord_flip()+
  facet_wrap(~param)+
  theme_paper()

plotname <- paste("biModality-",prefix, mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath,"figures/", plotname, sep=''), 
       width = 20, height = 30)

```

Write out expectations and MAPs of thetas (for Judith and future reference)

```{r}

d.stats <- d.params %>%
  filter(Parameter=="theta") %>%
  group_by(Item) %>%
  summarise(expectation = mean(Response),
            MAP = estimate_mode(Response),
            credHigh = HPDhi(Response),
            credLow = HPDlo(Response))



write.csv(d.stats, 
          file=paste('~/Documents/research/sinking-marbles/bayesian_model_comparison/priors/results/summary_', 
                     prefix, mhiter, "burn",  burn, ".csv", sep=''))

```


Global parameters (e.g. noise parameter)
```{r}
d.global<-d.params %>% filter(Item=='global')

ggplot(d.global, aes(x=Response))+
  geom_histogram(aes(y=..count../sum(..count..)), binwidth=0.01)+
  facet_wrap(~Parameter)+
  xlim(0,1)

plotname <- paste("phiPosterior-",prefix, mhiter/1000, 
                  "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath,"figures/", plotname, sep=''), 
       width = 8, height = 5)
```


By item CIs

```{r}

ggplot(d.tidy, aes(x=Item, y = MAP))+
  geom_bar(stat='identity', position=position_dodge(), alpha=0.4)+
  geom_errorbar(aes(ymin = credLow, ymax = credHigh))+
  facet_wrap(~Parameter)+
  theme(axis.text.x = element_text(angle=45))+
  coord_flip()



plotname <- paste("thetaCI-",prefix, mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath,"figures/", plotname, sep=''), 
       width = 20, height = 30)


qplot(data=d.tidy, x=MAP, geom='histogram')+
  facet_wrap(~Parameter)
d.tidy %>% filter(Parameter=='theta')
```

By item Posteriors

```{r}
ggplot(d.params %>% filter(Parameter=='theta'), aes(x=Response))+
  
ggplot(d.params %>% filter(Parameter=='theta'), aes(x=Response))+
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]))+ 
  facet_wrap(~Item)+
  xlim(0,1)+
  ylab("theta")







plotname <- paste("thetaPosteriors-",prefix, mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath, plotname, sep=''), 
       width = 30, height = 20)



ggplot(d.params %>% filter(Parameter=='phi'), aes(x=Response))+
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]))+ 
  facet_wrap(~Item)+
  xlim(0,1)+
  ylab("phi")

plotname <- paste("phiPosteriors-",prefix, mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath, plotname, sep=''), 
       width = 20, height = 20)
```


Compare to expected values of 4 steps
```{r}

to.n <- function(x) {return(as.numeric(as.character(x)))}
source("rscripts/helpers.r")


priorprobs = read.table(file="~/Documents/research/sinking-marbles/experiments/24_sinking-marbles-prior-fourstep/results/data/smoothed_15marbles_priors_withnames.txt",sep="\t", header=T, quote="")

#priorprobs = read.table(file="~/Documents/research/sinking-marbles/experiments/24_sinking-marbles-prior-fourstep/results/data/smoothed_15marbles_priors_withnames.txt",sep="\t", header=T, quote="")


row.names(priorprobs) = priorprobs$Item
nrow(priorprobs)

# prior all-state probability for each item
prior_allprobs = priorprobs[,c("Item","X15")]
row.names(prior_allprobs) = prior_allprobs$Item

# prior expectation for each item
gathered_probs <- priorprobs %>%
  gather(State,Probability,X0:X15)
gathered_probs$State = as.numeric(as.character(gsub("X","",gathered_probs$State)))
prior_exps <- gathered_probs %>%
  group_by(Item) %>%
  summarise(exp.val=sum(State*Probability))
prior_exps = as.data.frame(prior_exps)
row.names(prior_exps) = prior_exps$Item

bda.df<-left_join(d.tidy, prior_exps)

qplot(data=bda.df, y = MAP, x= exp.val, geom='point')+
  geom_errorbar(aes(ymin = credLow, ymax = credHigh))

```


From wonky-tfbt (should be the same)

```{r}

mhiter <- 50000
burn <- mhiter/2
samples <- mhiter - burn
fpath<- "~/Documents/research/sinking-marbles/bayesian_model_comparison/results/"
prefix <- "priorBDA-binomials_incrMH"
prefix <- "priorBDA-binomials-phi_incrMH"
#prefix <- "priorBDA-binomials-byItemPhi_incrMH"

d0<-read.csv(paste(fpath, prefix, mhiter, "burn",  burn, ".csv", sep=""))

d0 <- d0 %>% filter(Item!='global') %>% 
  select(-Parameter) %>%
  rename(Parameter = Item,
         Item = Value,
         Value = Probability)



d.params <- data.frame(Item = rep(d0$Parameter, 
                                         1+samples*d0$Value),
                         Estimate = rep(d0$Item, 
                                    1+samples*d0$Value))

d.params<- d.params %>% select(-Parameter) %>% rename(Estimate=Response)

d.tidy <- d.params %>% 
  group_by(Item) %>%
  summarise(MAP = estimate_mode(Estimate),
            credLow = HPDlo(Estimate),
            credHigh = HPDhi(Estimate))

d.binomials<- data.frame()
for (i in 1:length(d.tidy$Item)){
  theta <- d.tidy[i, "MAP"]$MAP
  print(i)
  d.binomials<- bind_rows(d.binomials,
                          data.frame(Item = d.tidy$Item[i],
                                    probs=dbinom(seq(0,15), 15, theta),
                                    bins = seq(0,15)
                                    )
                          )
  
}



```

Compare with emprical distributions of prior data

```{r}
emp.d0 <-  read.csv("~/Documents/research/sinking-marbles/bayesian_model_comparison/priors/data/priors.txt")
#emp.d0 <-  read.csv("~/Documents/research/sinking-marbles/bayesian_model_comparison/priors/data/priors_fourstep.txt")


empirical.Ns <- emp.d0 %>% 
  group_by(Item) %>%
  summarize(count = length(response))
  #summarize(count = length(best_guess))


#qplot(data=empirical.Ns, x = count, geom='histogram')


empirical.tidy <- emp.d0 %>%
  group_by(Item, response) %>%
  #group_by(Item, best_guess) %>%
  summarise(count = length(workerid)) %>%
  ungroup() %>%
  group_by(Item) %>%
  mutate(freq = count / sum(count)) %>%
  rename(bins = response)
  #rename(bins = best_guess)




fpath<- "~/Documents/research/sinking-marbles/bayesian_model_comparison/results/"
prefix<-"wonkyRSA_fullPosterior_unifBackoff_justCompState_wRepData_bdaPriors2thetas50k_so_wp_phi_incrMH"
#prefix <- "priorBDA-binomials-byItemPhi_incrMH"

d0<-read.csv(paste(fpath, prefix, mhiter, "burn",  burn, "a.csv", sep=""))

d.prior<-d0 %>% filter(Parameter=='prior') %>% 
  rename(bins = Quantifier) %>%
  group_by(Item, bins) %>%
  summarise(expval = sum(Value*Probability)) %>%
  ungroup() %>%
  mutate(bins = to.n(bins))



d.postpred <- d.params %>% 
  filter(!(Parameter%in%c('theta', "theta1", "theta2", "theta3",
                                             "mix", "w1", "w2", "w3", 'phi'))) %>% 
  group_by(Item, Parameter) %>%
  summarise(MAP = estimate_mode(Response),
            credLow = HPDlo(Response),
            credHigh = HPDhi(Response)) %>%
  ungroup() %>%
  rename(bins=Parameter) %>%
  mutate(bins = to.n(bins))


# map.predictive= data.frame()
# for (j in 1:length(d.tidy$Item)){
#     for (i in 0:15){
#            map.predictive<-bind_rows(map.predictive,
#                      data.frame(Item = d.tidy[[j, "Item"]],
#                       bins = i,
#                       MAPpredictive = dbinom(i, 15, d.tidy[[j, "MAP"]]),
#                       credLow = dbinom(i, 15, d.tidy[[j, "credLow"]]),
#                       credHigh = dbinom(i, 15, d.tidy[[j, "credHigh"]])
#                      ))
#     }
# }



#d.mixed<-left_join(map.predictive, empirical.tidy)
d.mixed<-left_join(d.postpred, empirical.tidy)
d.mixed<-left_join(d.prior, empirical.tidy)





d.mixed[is.na(d.mixed$count),]$freq <- 0


ggplot(d.mixed, aes(x=MAP, y = freq, color=factor(bins)))+
  
ggplot(d.mixed, aes(x=expval, y = freq, color=factor(bins)))+
  geom_point()+
  #geom_errorbarh(aes(xmin =credLow, xmax=credHigh))+
  geom_abline(intercept=0, slope=1, linetype = 2)+
  xlim(-0.05,1.05)+
  ylim(-0.05,1.05)+
  coord_fixed()+
  ggtitle(paste("r2 = ",round(with(d.mixed, cor(freq, expval))^2, digits=4),
                "\n MSE = ", round(with(d.mixed, mse(freq, expval)),digits=4), sep=''))


  ggtitle(paste("r2 = ",round(with(d.mixed, cor(freq, MAP))^2, digits=4),
                "\n MSE = ", round(with(d.mixed, mse(freq, MAP)),digits=4), sep=''))



plotname <- paste("scatter-modelData-intervals-",prefix, 
                  mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath,"figures/", plotname, sep=''), 
       width = 9, height = 9)
ggsave(file=paste(fpath,"graphs/model_curves/", plotname, sep=''), 
       width = 9, height = 9)



ggplot(d.mixed,
       aes(x=bins, y = freq))+
  geom_bar(position=position_dodge(), stat='identity')+
  geom_errorbar(aes(ymin=credLow, ymax=credHigh), position=position_dodge())+
  geom_bar(data=d.mixed,
                 aes(x=bins, 
                     y=MAP
                 ),
                 binwidth=1,
           fill='red',
           alpha=0.5,
                  stat='identity',
                 position=position_dodge())+
  facet_wrap(~Item)



plotname <- paste("barplot-modelData-",prefix, 
                  mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath,"figures/", plotname, sep=''), 
       width = 25, height = 20)


d.expectation <- d.postpred %>%
  group_by(Item) %>%
  summarise(exp.val = sum(to.n(bins)*MAP))

qplot(data=d.expectation, x=exp.val, geom='histogram')
range(d.expectation$exp.val)
plotname <- paste("expectationHistogram-",prefix, mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath,"figures/", plotname, sep=''), 
       width = 8, height = 5)
```




```{r}

ggplot(d.mixed %>% filter(freq > 0.25), aes(x=probs, y = freq))+
  geom_point()



d.mixed %>%
  group_by(bins) %>%
  summarise(cor(probs,freq))


with(d.mixed, cor(probs, freq))

ggplot(d.binomials,
       aes(x=bins, y = probs))+
  geom_bar(position=position_dodge(), stat='identity', alpha=0.5)+
  #geom_errorbar(aes(ymin=credLow, ymax=credHigh), position=position_dodge())+
  geom_histogram(data=emp.d0,
                 aes(x=response, 
                     y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]
                     ),
                 binwidth=1,
                 fill='blue',
                 alpha=0.3,
                 position=position_dodge())+
  facet_wrap(~Item)


ggplot(d.binomials,
       aes(x=bins, y = probs))+
  geom_bar(position=position_dodge(), stat='identity', alpha=0.5)+
  geom_histogram(data=cs.tidy %>% filter(Quantifier=='Some'),
                 aes(x=response, 
                     y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]
                     ),
                 binwidth=1,
                 fill='blue',
                 alpha=0.3,
                 position=position_dodge())+
  facet_wrap(~Item)



plotname <- paste("priorMAPpredsWData-",prefix, mhiter/1000, "kBurn", burn/1000, "k.pdf", sep='')
ggsave(file=paste(fpath, plotname, sep=''), 
       width = 35, height = 20)


ggplot(d.params, aes(x=Estimate))+
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]))+ 
  facet_wrap(~Item)+
  xlim(0,1)+
  ylab("theta")
```


Log likelihoods

```{r}

mhiter <- 50000
burn <- mhiter/2
samples <- mhiter - burn
fpath<- "~/Documents/research/sinking-marbles/bayesian_model_comparison/priors/results/"
prefix <- "priorBDA-logLikelihood-binomials_incrMH"
prefix <- "priorBDA-logLikelihood-binomials_phi-incrMH"
prefix <- "priorBDA-binomials-byItemPhi_incrMH"

d0<-read.csv(paste(fpath, prefix, mhiter, "burn",  burn, ".csv", sep=""))

d0 %>% summarise(sum(Value*Probability))


```

# Compare thetas with and without uniform guessing

```{r}

mhiter <- 50000
burn <- mhiter/2
samples <- mhiter - burn
fpath<- "~/Documents/research/sinking-marbles/bayesian_model_comparison/priors/results/"

prefix <- "priorBDA-binomials-phi-fullPosterior_incrMH"
d0<-read.csv(paste(fpath, prefix, mhiter, "burn",  burn, ".csv", sep=""))



d.params <- data.frame(Parameter = rep(d0$Parameter, 
                                         1+samples*d0$Probability),
                         Item = rep(d0$Item, 
                                    1+samples*d0$Probability),
                         Response = rep(d0$Value, 
                                        1+samples*d0$Probability))


d.tidy <- d.params %>% filter(Parameter=='theta') %>% 
  group_by(Parameter,Item) %>%
  summarise(MAP = estimate_mode(Response),
            credLow = HPDlo(Response),
            credHigh = HPDhi(Response))

qplot(data=d.tidy, x=MAP, geom='histogram')

prefix <- "priorBDA-binomials-fullPosterior_incrMH"
d0<-read.csv(paste(fpath, prefix, mhiter, "burn",  burn, ".csv", sep=""))



d.params <- data.frame(Parameter = rep(d0$Parameter, 
                                         1+samples*d0$Probability),
                         Item = rep(d0$Item, 
                                    1+samples*d0$Probability),
                         Response = rep(d0$Value, 
                                        1+samples*d0$Probability))


d.tidy2 <- d.params %>% filter(Parameter=='theta') %>% 
  group_by(Parameter,Item) %>%
  summarise(MAPsansPhi = estimate_mode(Response),
            credLowsansPhi = HPDlo(Response),
            credHighsansPhi = HPDhi(Response))

qplot(data=d.tidy2, x=MAPsansPhi, geom='histogram')

d.tidy3<-left_join(d.tidy,d.tidy2)

d.tidy3<-d.tidy3 %>% mutate(diff = abs(MAP-MAPsansPhi))
d.tidy3 <- d.tidy3[with(d.tidy3,order(diff)),]
```



